### 2012-07-22 developing mind/proposal developing mind.tex
proposal finished draft check conclusion

\input{$HOME/Documents/submissions/preamble_steve_report}
%\author{}

@@

\section{Competition}
At the time of writing there is no book or collection devoted to philosophical issues in cognitive development other than \textit{The Philosophical Baby} \citep{gopnik:2009_philosophical} which is primarily aimed at general readers.

@@

\subsection{Introduction}
How do humans come to know  about---and to knowingly manipulate---%
objects,
causes,
words,
numbers,
colours,
actions
and
minds?
In pursuing this question we have to consider minds where the knowledge is neither clearly present nor obviously absent. 
This is challenging because both commonsense and theoretical tools for describing minds are generally suitable only for fully developed adult minds. 
Davidson writes:
%
\begin{quote}
`%
%if you want to describe what is going on in the head of the child when it has a few words which it utters in appropriate situations, you will fail for lack of the right sort of words of your own. 
We have many vocabularies for describing nature when we regard it as mindless, and we have a mentalistic vocabulary for describing thought and intentional action; what we lack is a way of describing what is in between' \citep[p.\ 11]{Davidson:1999ju}.
\end{quote}
%
%Of course you might not agree with Davidson that children just starting to use words are incapable of thought or of intentional action. 
%In fact research on children's pre-verbal communicative behaviours such as pointing seems not only to manifest purposiveness but also a sensitivity to others' thoughts.
%(Davidson seems to have paid little attention to development, and he held that being able to think, to act intentionally and to communicate with language all require understanding the possibility that one's own beliefs may be false.)
%Even so, Davidson 
%
To understand the emergence of knowledge we need to find ways of describing what is in between: individuals whose movements are neither mindless nor guided by intention and knowledge.  
Some progress has already been made but many challenges remain. 
Philosophers have much to contribute in crafting distinctions and conceptual tools useful for meeting the challenges involved in describing what is in between. 
There is also much for philosophers to gain: studying what is in between mindless nature and the sorts of cognition captured by   commonsense psychological notions 
 will reveal things about what minds are and how they work.

@@

\subsection{Chapter 5. Words and Other Communicative Tools}
One problem in acquiring a language is to identify which things are words; among much else, this involves working out where, in a stream of speech or gesture, one word ends and another begins.  
In solving this problem infants may rely on statistical learning \citep{Saffran:1996aj}.
Another problem is learning which words pick out which objects and properties; this is often called the `mapping problem'. 
Philosophers have sometimes held that in solving this problem children rely on association (or perhaps other forms of statistical learning).
Compare Wittgenstein (\citeyear[p.\ 77]{Wittgenstein:1972lj}): 
%
\begin{quote}
`The child learns this language from the grown-ups by being trained to its use. I am using the word ``trained'' in a way strictly analogous to that in which we talk of an animal being trained to do certain things. It is done by means of example, reward, punishment, and suchlike.'
%careful with this quote --- from the context Wittgenstein is imagining how things might be, not how they are.
%but this quote shows that he is serious about how children are:
%`A child uses such primitive forms of language when it learns to talk. Here the teaching of language is not explanation, but training.' \citep[§5, p.\ 4]{Wittgenstein:1953mm}
%And:
%`The primitive language-game which children are taught needs no justification; attempts at justification need to be rejected.' \citep[p.\ 200]{Wittgenstein:1953mm}
%And:
%`I myself wrote in my book that children learn to understand a word in such and such a way. Do I know that, or do I believe it? Why in such a case do I write not "I believe etc." but simply the indicative sentence?' \citep[§290, p.\ 38]{Wittgenstein:1974dk}
%
%See also Russell (1921: 71): “A child learning to speak is learning habits and associations which are just as much determined by the environment as the habit of expecting dogs to bark and cocks to crow. The community that speaks a language has learnt it, and modified it by processes almost all of which are not deliberate, but the results of causes operating according to more or less ascertainable laws.”
\end{quote}
%
If this sort of view were right, abilities to communicate linguistically would not necessarily depend on knowledge proper and so it would be possible to appeal to linguistic communication in explaining the developmental origins of knowledge. 
%Compare Wittgenstein again:
%`535. The child knows what something is called if he can reply correctly to the question "what is that called?"
%536. Naturally, the child who is just learning to speak has not yet got the concept is called at all.
%537. Can one say of someone who hasn't this concept that he knows what such-and-such is called?
%538. The child, I should like to say, learns to react in such-and-such a way; and in so reacting it doesn't so far know anything. Knowing only begins at a later level.'
%\citep[p.\ 71]{Wittgenstein:1974dk}
This does indeed seem plausible for the special case of colour words, as we saw in Chapter 4 (on Colours); 
and something like it may hold for number words too (see Chapter 3 on Numbers).  
%ref to Wittgenstein on child learning about numbers somewhere?
But putting these special cases aside, the experimental findings do not support Wittgenstein's view.
Instead they support the hypothesis that children solve the mapping problem by general reasoning and rely on social cognition
\citep{Baldwin:1995xl,Bloom:2000qz,Sabbagh:2001sp}. 
%Also talk about language creation here (see lexical innovation paper)

This hypothesis conflicts with Davidson's claims that having thoughts involves grasping the concept of truth \citep[p.\ 189]{Davidson:2001sm}, 
% `What gives my belief its content ... is my knowledge of what is required for the belief ... to be true' \citep[p.\ 189]{Davidson:2001sm}, 
and that `we grasp the concept of truth only when we can communicate the contents—the propositional contents—of the shared experience, and this requires language' \citep[p.\ 27]{Davidson:1997wj}.
% also:
%`language ... is, or has become, a mode of perception.  However, speech is not just one more organ; it is essential to the other senses if they are to yield propositional knowledge' \citep[p.\ 22]{Davidson:1997wj}
% also:
%`Mastery of the word or concept [`table'] requires in addition, I believe, that the child understand that error is possible, that "That's a table' expresses a judgment that has a truth value independent of its merely being uttered.' \citep[p.\ 9]{Davidson:2001np}
%
The objection to Davidson's view is that  it has things backwards: acquiring your first words already involves thinking about what they pick out. 
If this objection is correct, can anything be salvaged from Davidson's arguments? 
While the objection rules out claiming that being able to think depends on being able to communicate propositional contents using language, it leaves open the possibility that thought is essentially intersubjective.
One challenge, then, is to explain the intersubjectivity of thought without appeal to communication by language.

@@

Tracking a goal-directed action involves at least two things: distinguishing the action from other events 
and
identifying to which goal or goals it is directed.
Infants' expectations indicate that they can do both of these things from some time around the end of their first year of life \citep[e.g.][]{Csibra:2003kp}. 
But tracking is not understanding.
To say that infants can track goal-directed action is not yet to say anything about what they understand of action.
Consider two questions about the nature of action.
First, which events are actions?
Second, what is the relation between an action and the goal or goals to which the action is directed?
For each of these we can ask corresponding questions about infants' understanding of action.
Which events do infants take to be actions?
And how do they understand the relation between actions and the goals to which they are directed?
We need to answer these questions in order to understand how early abilities to track action are developmentally related to knowledge of action, and also how they relate to developments in other domains including language.
In this chapter we consider what infants understand of action.

@@

Are there alternative views of action understanding which do not demand knowledge of intentions?
As before, it is helpful to start with a view about what actions are and then consider corresponding views about understanding actions.
Instead of thinking of the relation between actions and the goals to which they are directed as determined by intentions representing those goals,
we might instead attempt to explain the relation in terms of justification.
This is the idea behind Gergely and Csibra's `principle of rational action' which states that:
%
\begin{quote}
`%the principle of rational action ... states that 
an action can be explained by a goal state if, and only if, it is seen as the most justifiable action towards that goal state that is available within the constraints of reality' 
(Csibra and Gergely 1998: 255; cf. Csibra, Bíró, et al. 2003)
\end{quote}
%
%*Don't we have to separate out the question of which events are actions?
This view of action understanding requires less conceptual sophistication thanks to not requiring knowledge of intentions as such. 
But remains quite demanding, for it entails that one-year-olds  are able to justify actions and engage in means end reasoning.
Gergely and Csibra stress this aspect of their view:
%
\begin{quote}
`%when taking the teleological stance 
one-year-olds apply the same inferential principle of rational action that drives everyday mentalistic reasoning about intentional actions in adults' (\citealp[p.\ 259]{Csibra:1998cx}; \citealp[compare][p.\ 290]{Gergely:2003gb})
\end{quote}
%
This is incompatible with the notion that infants' abilities to track action depend on core knowledge or modular representations.
For, as we saw in Chapter 2 (on Objects),  processes involving core knowledge  are supposed to be informationally encapsulated;
and informationally encapsulated processes cannot represent relations involving rationality or justification as such \citep{Fodor:2000cj}.
To see whether it is even theoretically coherent to suppose that understanding action could involve core knowledge only, we must find alternative accounts of what it is to understand actions, accounts that involve neither intention nor rationality.

@@

So far we have ignored the possibility that motor representations and processes may play a role in understanding action.
It is now well established that motor processes and representations are involved in observing actions---not only in producing them \citep{rizzolatti_functional_2010}.
Furthermore, some motor representations represent outcomes to which actions are directed \citep{jeannerod_motor_2006}. 
For instance, observing another grasping a cup can involve a motor representation of the outcome, of the cup's being grasped, and not only motor representations of merely kinematic or dynamic features.
Apparently, then, motor planning can be used in reverse: it can be used for inferring a goal from an observed action and not only for  determining what to do in order to achieve a particular goal.
We also know that, in adults and infants alike, motor representation can facilitate judgements about the goals of actions 
(on adults: \citealp{serino:2009_lesions_,pazzaglia:2008_sound_}; on infants: \citealp{bekkering:2000_imitation}).
It is possible, therefore, that infants can track actions not because they know about agent's intentions or consider the rationality of actions,
but because their motor expertise enables  them to  determine to which goals others' actions could be directed. 
(This is surely not the whole story, but it can be extended to involve statistical learning as well as motor representation; see \citet{paulus:2011_role}, \citet{Baldwin:2001rn}, and \citet{Saylor:2007pj}.)

@@

\subsection{Conclusion}
Are we closer to understanding 
how humans come to know about %
objects,
causes,
words,
numbers,
colours,
actions
and
minds? 
***

>>>>

%\input{$HOME/Documents/submissions/preamble_steve_report}
\input{$HOME/Documents/submissions/preamble_steve_paper2}

%adjust title size & font
\pretitle{
	\begin{center}
	%\sffamily
	\LARGE
}

%\author{}

%change fonts (not anonymous)
%\setromanfont[Mapping=tex-text]{Linux Libertine O} 
%\setsansfont[Mapping=tex-text]{Linux Biolinum O} 
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

@@

\section{Competition}
At the time of writing there is no book or collection devoted to philosophical issues in cognitive development other than \textit{The Philosophical Baby} \citep{gopnik:2009_philosophical}  is primarily aimed at general readers.

@@

\subsection{Introduction}
How do humans come to know  about---and to knowingly manipulate---%
objects,
causes,
words,
numbers,
colours,
actions
and
minds?
In pursuing this question we have to consider minds where the knowledge is neither clearly present nor obviously absent. 
This is challenging because both commonsense and theoretical tools for describing minds are generally suitable only for fully developed adult minds. 
Davidson writes:
%
\begin{quote}
`%
%if you want to describe what is going on in the head of the child when it has a few words which it utters in appropriate situations, you will fail for lack of the right sort of words of your own. 
We have many vocabularies for describing nature when we regard it as mindless, and we have a mentalistic vocabulary for describing thought and intentional action; what we lack is a way of describing what is in between' \citep[p.\ 11]{Davidson:1999ju}.
\end{quote}
%
%Of course you might not agree with Davidson that children just starting to use words are incapable of thought or of intentional action. 
%In fact research on children's pre-verbal communicative behaviours such as pointing seems not only to manifest purposiveness but also a sensitivity to others' thoughts.
%(Davidson seems to have paid little attention to development, and he held that being able to think, to act intentionally and to communicate with language all require understanding the possibility that one's own beliefs may be false.)
%Even so, Davidson 
%
To understand the emergence of knowledge we need to find ways of describing what is in between: individuals whose movements are neither mindless nor guided by intention and knowledge.  
And the ways we find to describe these individuals 
%that are not only theoretically coherent and empirically motivated---in addition,
%we need ways of describing these individuals 
must enable us to understand what it is like to be them, to
see things from their points of view.
Some progress has already been made but many challenges remain. 
Philosophers have much to contribute in crafting distinctions and conceptual tools useful for meeting the challenges involved in describing what is in between. 
There is also much for philosophers to gain: studying what is in between mindless nature and the sorts of cognition captured by   commonsense psychological notions 
 will reveal things about what minds are and how they work.

@@

\subsection{Chapter 5. Words and Other Communicative Tools}
One problem in acquiring a language is to identify which things are words; among much else, this involves working out where, in a stream of speech or gesture, one word ends and another begins.  
In solving this problem infants may rely on statistical learning \citep{Saffran:1996aj}.
Another problem is learning which words pick out which objects and properties; this is often called the `mapping problem'. 
Philosophers have sometimes held that in solving this problem children rely on association (or perhaps other forms of statistical learning).
Compare Wittgenstein (\citeyear[p.\ 77]{Wittgenstein:1972lj}): 
%
\begin{quote}
`The child learns this language from the grown-ups by being trained to its use. I am using the word ``trained'' in a way strictly analogous to that in which we talk of an animal being trained to do certain things. It is done by means of example, reward, punishment, and suchlike.'
%careful with this quote --- from the context Wittgenstein is imagining how things might be, not how they are.
%but this quote shows that he is serious about how children are:
%`A child uses such primitive forms of language when it learns to talk. Here the teaching of language is not explanation, but training.' \citep[§5, p.\ 4]{Wittgenstein:1953mm}
%And:
%`The primitive language-game which children are taught needs no justification; attempts at justification need to be rejected.' \citep[p.\ 200]{Wittgenstein:1953mm}
%And:
%`I myself wrote in my book that children learn to understand a word in such and such a way. Do I know that, or do I believe it? Why in such a case do I write not "I believe etc." but simply the indicative sentence?' \citep[§290, p.\ 38]{Wittgenstein:1974dk}
%
%See also Russell (1921: 71): “A child learning to speak is learning habits and associations which are just as much determined by the environment as the habit of expecting dogs to bark and cocks to crow. The community that speaks a language has learnt it, and modified it by processes almost all of which are not deliberate, but the results of causes operating according to more or less ascertainable laws.”
\end{quote}
%
If this sort of view were right, abilities to communicate linguistically would not necessarily depend on knowledge proper. 
So it would be possible to conjecture that the emergence in development of knowledge proper is a consequence of 
training with language.
%Compare Wittgenstein again:
%`535. The child knows what something is called if he can reply correctly to the question "what is that called?"
%536. Naturally, the child who is just learning to speak has not yet got the concept is called at all.
%537. Can one say of someone who hasn't this concept that he knows what such-and-such is called?
%538. The child, I should like to say, learns to react in such-and-such a way; and in so reacting it doesn't so far know anything. Knowing only begins at a later level.'
%\citep[p.\ 71]{Wittgenstein:1974dk}
This does indeed seem plausible for the special case of colour words, as we saw in Chapter 4 (on Colours); 
and something like it may hold for number words too (see Chapter 3 on Numbers).  
%ref to Wittgenstein on child learning about numbers somewhere?
But putting these special cases aside, the experimental findings do not support Wittgenstein's claim that language is entirely acquired through training.
Instead they support the hypothesis that children solve the mapping problem by general reasoning and rely on social cognition
\citep{Baldwin:1995xl,Bloom:2000qz,Sabbagh:2001sp}. 
This hypothesis is further supported by findings that children who have learnt few or no words sometimes invent their own words \citep{Clark:1993bv}, 
and by
research on profoundly deaf children brought up in purely oral environments and therefore without experience of language who, individually or in groups, create their own signed languages \citep{Kegl:1999es,Senghas:2001zm,Goldin-Meadow:2003pj}.

This hypothesis---children solve the mapping problem by general reasoning---conflicts with Davidson's claims that having thoughts involves grasping the concept of truth \citep[p.\ 189]{Davidson:2001sm}, 
% `What gives my belief its content ... is my knowledge of what is required for the belief ... to be true' \citep[p.\ 189]{Davidson:2001sm}, 
and that `we grasp the concept of truth only when we can communicate the contents—the propositional contents—of the shared experience, and this requires language' \citep[p.\ 27]{Davidson:1997wj}.
% also:
%`language ... is, or has become, a mode of perception.  However, speech is not just one more organ; it is essential to the other senses if they are to yield propositional knowledge' \citep[p.\ 22]{Davidson:1997wj}
% also:
%`Mastery of the word or concept [`table'] requires in addition, I believe, that the child understand that error is possible, that "That's a table' expresses a judgment that has a truth value independent of its merely being uttered.' \citep[p.\ 9]{Davidson:2001np}
%
The objection to Davidson's view is that  it has things backwards: acquiring your first words already involves thinking about what they pick out. 
If this objection is correct, can anything be salvaged from Davidson's arguments? 
While the objection rules out claiming that being able to think depends on being able to communicate propositional contents using language, it leaves open the possibility that thought is essentially intersubjective.
One challenge, then, is to explain the intersubjectivity of thought without appeal to communication by language.

@@

Tracking a goal-directed action involves at least two things: distinguishing the action from other events 
and
identifying to which goal or goals it is directed.
Infants' expectations indicate that they can do both of these things from some time around the end of their first year of life \citep[e.g.][]{Csibra:2003kp}. 
But tracking is not understanding.
To say that infants can track goal-directed action is not yet to say anything about what they understand of action.
Consider two questions about the nature of action.
First, which events are actions?
Second, what is the relation between an action and the goal or goals to which the action is directed?
For each of these we can ask corresponding questions about infants' understanding of action.
Which events do infants take to be actions?
And how do they understand the relation between actions and the goals to which they are directed?
We need to answer these questions in order to understand how early abilities to track action are developmentally related to knowledge of action, and also how they relate to developments in other domains including language.
In this chapter we explore three hypotheses about the understanding which underpins infants' abilities to track action.

@@

Are there alternative views of what understanding actions involves which do not demand knowledge of intentions?
As before, it is helpful to start with a view about what actions are and then consider corresponding views about understanding actions.
Instead of thinking of the relation between actions and the goals to which they are directed as determined by intentions representing those goals,
we might instead attempt to explain the relation in terms of justification.
This is the idea behind Gergely and Csibra's `principle of rational action' which states that:
%
\begin{quote}
`%the principle of rational action ... states that 
an action can be explained by a goal state if, and only if, it is seen as the most justifiable action towards that goal state that is available within the constraints of reality' 
(Csibra and Gergely 1998: 255; cf. Csibra, Bíró, et al. 2003)
\end{quote}
%
%*Don't we have to separate out the question of which events are actions?
As part of an account of what actions are, this is unlikely to be correct.
After all, agents sometimes fail to select actions that could be justified (or seen as justifiable) given the goal to which they are directed.
\citet[p.\ 123]{Csibra:2003jv} themselves note this point.
But the principle of rationality might serve as a useful approximation.
Perhaps someone who applied the principle of rationality to work out which goals actions are directed to would be right much of the time. 
So perhaps infants' abilities to track actions involve understanding action in accordance with the principle of rationality.
This view of action understanding requires less conceptual sophistication thanks to not requiring knowledge of intentions or other propositional attitudes as such. 
But it remains quite demanding, for it entails that one-year-olds  are able to justify actions and engage in means end reasoning.
Gergely and Csibra stress this aspect of their view:
%
\begin{quote}
`%when taking the teleological stance 
one-year-olds apply the same inferential principle of rational action that drives everyday mentalistic reasoning about intentional actions in adults' (\citealp[p.\ 259]{Csibra:1998cx}; \citealp[compare][p.\ 290]{Gergely:2003gb})
\end{quote}
%
This is incompatible with the notion that infants' abilities to track action depend on core knowledge or modular representations.
For, as we saw in Chapter 2 (on Objects),  processes involving core knowledge  are supposed to be informationally encapsulated;
and informationally encapsulated processes cannot represent relations involving rationality or justification as such \citep{Fodor:2000cj}.
To see whether it is even theoretically coherent to suppose that understanding action could involve core knowledge only, we must find alternative accounts of what it is to understand actions, accounts that involve neither intention nor rationality.

@@

So far we have ignored the possibility that motor representations and processes play a role in understanding action.
It is now well established that motor processes and representations are involved in observing actions---not only in producing them \citep{rizzolatti_functional_2010}.
Furthermore, some motor representations represent outcomes to which actions are directed \citep{jeannerod_motor_2006}. 
For instance, observing another grasping a cup can involve a motor representation of the outcome, of the cup's being grasped, and not only motor representations of merely kinematic or dynamic features.
Apparently, then, motor planning can be used in reverse: it can be used for inferring a goal from an observed action and not only for  determining what to do in order to achieve a particular goal.
We also know that, in adults and infants alike, motor representation can facilitate judgements about the goals of actions 
(on adults: \citealp{serino:2009_lesions_,pazzaglia:2008_sound_}; on infants: \citealp{bekkering:2000_imitation}).
It is possible, therefore, that infants can track actions not because they know about agent's intentions or consider the rationality of actions,
but because their motor expertise enables  them to  determine to which goals others' actions could be directed. 
(This is surely not the whole story, but it can be extended to involve statistical learning as well as motor representation; see \citet{paulus:2011_role}.)
%don't cite here: these are to do with segmentation
%, \citet{Baldwin:2001rn}, and \citet{Saylor:2007pj}.)
On this hypothesis, infants' abilities to track action do not involve understanding actions in the way that human adults do (or are standardly held to). 
For this reason the hypothesis might  appear  to make infants' abilities concerning action philosophically uninteresting.
We can see that this appearance is mistaken by reflecting on how intentions (and other propositional attitudes) might refer to actions involving bodily movements.
The considerations for the view that 
thought about objects depends on perceptual experience of them  \citep[see][]{Campbell:2002ge} 
also support the view that 
all thought about actions ultimately depends on motor experience of them.

@@

\subsection{Conclusion}
Are we closer to understanding 
how humans come to know about %
objects,
causes,
words,
numbers,
colours,
actions
and
minds? 
Maybe.
%In some special cases---number, colour, words and minds---there are theories which make testable predictions.
%These theories conflict with some philosophers' views about development.
%Most obviously, these theories provide objections to any form of nativism on which number, colour or psychological concepts are innate.
%The findings that inform these theories also lead to objections to ***Wittgenstein ***Davidson
We have seen that apparent inconsistencies in developmental findings concerning knowledge of objects (Chapter 2) and knowledge of minds (Chapter 7) can be resolved by distinguishing kinds of knowledge.
One challenge was to make such distinctions in ways that are both  theoretically coherent and empirically motivated.
A first step was to appeal to the notions of modularity and core knowledge; a further step was to consider how, in particular cases (including objects, numbers, colours and actions), perceptual and motor representations and processes can constitute core knowledge.

A second, related challenge arises from hypotheses that social interaction enables the emergence in development of knowledge.  
For this hypothesis to be coherent, there must be forms of social interaction which do not already presuppose the knowledge whose emergence is to be explained.
Existing accounts of social interaction apply only to those who already have knowledge of several domains, including objects, actions and minds.
The challenge, then, is to characterise these simple forms of social interaction.

Distinguishing core knowledge from knowledge proper 
and
characterising simple forms of social interaction
are 
two key theoretical ingredients of
developmental theories with testable predictions for some special cases, including numbers, colours, words and minds.
These theories will probably turn out to be wrong, of course.
What seems secure, though, is the idea that explaining development requires 
%1
distinguishing multiple kinds of knowledge, 
%2
identifying simple forms of social interaction 
and 
%3
understanding how 
	these jointly enable development.
	%social interaction can transform knowledge.
